%\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{Bibliography}


\begin{thebibliography}{99}

\bibitem{fatamorgana}
Fata Morgana: \textit{http://www.fatamorgana.ch/tichu/tichu.asp}

\bibitem{rules}
Tichu Rules and strategies: \textit{http://scv.bu.edu/$\sim$aarondf/Games/Tichu/}

\bibitem{pimcts}
Jeffrey Long, Nathan R. Sturtevant, Michael Buro, Timothy Furtak,
\textbf{Understanding the Success of Perfect Information Monte Carlo Sampling in Game Tree Search},
Department of Computing Science, University of Alberta Edmonton, Alberta, Canada
in Proceedings of the Twenty-Fourth AAAI Conference on Artificial Intelligence, 2010

\bibitem{david12}
David Pfander,
\textbf{Eine künstliche Intelligenz für das Kartenspiel Tichu},
Universität Stuttgart,
2013.

\bibitem{uct}
L. Kocsis and C. Szepesvári, \textbf{Bandit based Monte-Carlo Planning} in Euro. Conf. Mach. Learn. Berlin, Germany: Springer, 2006, pp. 282–293.

\bibitem{bridge}
I. Frank and D. Basin, \textbf{Search in games with incomplete information: A case study using Bridge card play}
Artif. Intell., vol. 100, no. 1–2,
pp. 87–123, 1998.

\bibitem{ismcts}
Peter I. Cowling, Edward J. Powley, Daniel Whitehouse,
\textbf{Information Set Monte Carlo Tree Search},
IEEE TRANSACTIONS ON COMPUTATIONAL INTELLIGENCE AND AI IN GAMES, VOL. 4, NO. 2,
JUNE 2012

\bibitem{perez13}
Diego Perez, Spyridon Samotharakis, Simon M.Lucas,
\textbf{Rolling Horizon Evolution versus Tree Search for Navigation in Single-Player Real-Time Games},
Juli 2013

\bibitem{russel14}
Stuart Russel, Peter Norvig
\textbf{Artificial Intelligence, a modern approach},
third edition,
2014.

\bibitem{surveymcts}
Cameron Browne, Edward Powley, Daniel Whitehouse, Simon Lucas, Peter I. Cowling, Philipp Rohlfshagen, Stephen Tavener, Diego Perez, Spyridon Samothrakis and Simon Colton,
\textbf{A Survey of Monte Carlo Tree Search Methods},
IEEE TRANSACTIONS ON COMPUTATIONAL INTELLIGENCE AND AI IN GAMES, VOL. 4, NO. 1,
MARCH 2012



\bibitem{whitehouse14}
Daniel Whitehouse,
\textbf{Monte Carlo Tree Search for games with Hidden Information and Uncertainty},
University of York,
August 2014

\bibitem{movegroups}
Benjamin E. Childs, James H. Brodeur,
\textbf{Transpositions and Move Groups in Monte Carlo Tree Search},

\bibitem{movegroups2}
Gabriel Van Eyck, Martin Müller,
\textbf{Revisiting Move Groups in Monte Carlo Tree Search},
University of Alberta, Edmonton, Canada,
January 2012

\bibitem{7wonders}
D. Robilliard, C. Fonlupt, and F. Teytaud
\textbf{Monte-Carlo Tree Search for the Game of “7 Wonders”},
LISIC, ULCO, Univ Lille–Nord de France, FRANCE


\bibitem{tichumania}
Tichu game-logs provided by \textit{http://www.brettspielwelt.de/}: \textit{http://log.tichumania.de}

\bibitem{alphago}
Silver and Huang et al, \textbf{Mastering the game of Go with deep neural networks and tree search},
Google DeepMind, nature vol 529, p484+, January 2016

\bibitem{kerasrl}
Github page of keras-rl python package: \textbf{https://github.com/matthiasplappert/keras-rl}

% TODO read:
%\bibitem{Frank98}
%I. Frank and D. Basin, \textbf{Search in games with incomplete information:
%A case study using Bridge card play} Artif. Intell., vol. 100, no. 1–2,
%pp. 87–123, 1998.



\end{thebibliography}
